---
title: "PML"
author: "Ankur"
date: "22/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
getwd()
setwd("C:/Users/ankur/Documents")

```
#SETUP

## Due to sample of the training sample Parallel Processing was implemented
```{r}

library(doParallel)
suppressWarnings(suppressMessages(library(caret)))
suppressWarnings(suppressMessages(library(randomForest)))
suppressWarnings(suppressMessages(library(e1071)))
set.seed(0007)

```

# INPUT

## Download source data
```{r}
trainingFilename   <- 'pml-training.csv'
quizFilename       <- 'pml-testing.csv'
trainingUrl        <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
quizUrl            <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

```

## Data Cleansing

### Remove NA and Blank Values with na.strings parameter
```{r}
Training_1     <-read.csv(trainingFilename, na.strings=c("NA","","#DIV/0!"))
Training_1     <-Training_1[,colSums(is.na(Training_1)) == 0]
dim(Training_1)

```

```{r}
quiz.df         <-read.csv(quizFilename , na.strings=c("NA", "", "#DIV/0!"))
quiz.df         <-quiz.df[,colSums(is.na(quiz.df)) == 0]
dim(quiz.df)

```
## Reduce the number of variables
```{r}
Training_1   <-Training_1[,-c(1:7)]
Quiz.df <-quiz.df[,-c(1:7)]
dim(Training_1)

```
## Check for near zero values in training data
```{r}

Training.nzv<-nzv(Training_1[,-ncol(Training_1)],saveMetrics=TRUE)

```
## Display and count variables submitted for the train function
```{r}

rownames(Training.nzv)

dim(Training.nzv)[1]

```
# ALGORITHM

## Partition the training data into a training set and a testing/validation set
```{r}

inTrain     <- createDataPartition(Training_1$classe, p = 0.6, list = FALSE)
inTraining  <- Training_1[inTrain,]
in_test      <- Training_1[-inTrain,]
dim(inTraining);dim(in_test)

```
## Construct the model using cross validation
```{r}

myModelFilename <- "myModel.RData"
if (!file.exists(myModelFilename)) {

    # Parallel cores  
    #require(parallel)
    library(doParallel)
    ncores <- makeCluster(detectCores() - 1)
    registerDoParallel(cores=ncores)
    getDoParWorkers() # 3    
    
    # use Random Forest method with Cross Validation, 4 folds
    myModel <- train(classe ~ .
                , data = inTraining
                , method = "rf"
                , metric = "Accuracy"  # categorical outcome variable so choose accuracy
                , preProcess=c("center", "scale") # attempt to improve accuracy by normalising
                , trControl=trainControl(method = "cv"
                                        , number = 4 # folds of the training data
                                        , p= 0.60
                                        , allowParallel = TRUE 
#                                       , seeds=NA # don't let workers set seed 
                                        )
                )

    save(myModel, file = "myModel.RData")
    # 3:42 .. 3:49 without preProcess
    # 3:51 .. 3:58 with preProcess
    stopCluster(ncores)
} else {
    # Use cached model  
    load(file = myModelFilename, verbose = TRUE)
}

```

```{r}

print(myModel, digits=4)

```
# PREDICT
```{r}

predTest <- predict(myModel, newdata=in_test)

```
## Out of Sample Error
### Accuracy is .9963 and this figure lies within the 95% confidence interval

# FINAL MODEL DATA

```{r}

myModel$finalModel

```

```{r}

varImp(myModel)

```
# VALIDATION
```{r}

print(predict(myModel, newdata=Quiz.df))

```

